---
title: "In-Class Exercise 5-1"
author: "Aaron Melcher"
output: html_document
date: "`r Sys.Date()`"
runtime: shiny
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
# Setup with initial packages
library(tidyverse)
library(shiny)
library(DT) # for shiny's table
theme_set(theme_bw())

# Read in data
birds <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/exercises/data/birds.csv") |>
    separate(Species, c("genus", "species2"))

# Create a scatter function for use in the ui page
scatter <- function(df, selected) {
    palette <- c("#d4d4d4", "black")
    df |>
        mutate(selected_ = selected) |>
        ggplot() +
        geom_point(aes(Asymmetry, Ellipticity, color = as.factor(selected_))) +
        scale_color_manual(values = palette)
}

# Generate the Shiny UI basics
ui <- fluidPage(
  titlePanel("Birds Brushed Scatterplot"),
  plotOutput("plot", brush = "scatter_brush"),
  DTOutput("table")
)

# Server function for use in the shiny ui
server <- function(input, output) {
    selected <- reactiveVal(rep(TRUE, nrow(birds)))
     observeEvent(
        input$scatter_brush, {
        brushed_points <- brushedPoints(birds, input$scatter_brush, allRows = TRUE)
        selected(brushed_points$selected_)
    }
  )
  output$plot <- renderPlot(scatter(birds, selected()))
  output$table <- renderDT(filter(birds, selected()))
}

# Start the Shiny server
shinyApp(ui, server)
```

The code above creates an interactive graph that allows for brushing to select a subset of the graphed data to see the information provided. The strategy is to first read in the data and split it on genus and species2 for further use. Then a scatter function is defined as well as the UI and Server functions to be used in the Shiny UI. The UI defines the basics of the graph like name and what plot are to be used. Server defines the events that can happen on the Shiny UI such as the brushing and also how the table should interact with the brushing.